Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
writing

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Tbl.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.

%% The neural recognition model accelerates inference,
%% but given enough iterations of our algorithm,
%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Fig.~\ref{regressionLearningCurve}; No NN in Tbl.~\ref{regressionComparison}).
%% We compare our model on held-out tasks against 4 baselines (Tbl.~\ref{regressionComparison}).

\begin{figure*}[h]
\makebox[\textwidth][c]{
\begin{minipage}{0.55\textwidth} \newcommand{\functionSize}{1.5cm}\centering
%  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
  %% \includegraphics[width = \functionSize]{figures/functions/432.png}
  %% \includegraphics[width = \functionSize]{figures/functions/415.png}
  %% \includegraphics[width = \functionSize]{figures/functions/426.png}
  %% \includegraphics[width = \functionSize]{figures/functions/597.png}
  %% \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{minipage}
\hspace{1em}
\begin{minipage}{0.5\textwidth}
  \centering\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
  \\\bottomrule
\end{tabular}
  \captionof{table}{Some learned subroutines for symbolic regression. System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{minipage}
}
\end{figure*}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

@@

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}}& RF/DC & PCFG & Enum

@@

  \\\midrule\multicolumn{6}{c}{\emph{List Processing}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s

  \\\midrule\multicolumn{6}{c}{\emph{Text Editing}}\\\midrule
  \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\
  Solve time&  29s&s&80s&--&235s

@@

  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&   \textbf{84\% }&75\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &31s&55s&29s
  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6

  \\\bottomrule
  \end{tabular}
\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}

>>>>

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks. The learned DSL contains \textbf{X} new functions, listed in the supplement.

@@

\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
The algorithm also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{r}{0.45\textwidth} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
    \end{wrapfigure}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{EC}, which lesions the recognition model and modifies  the DSL learner to model the library learning technique of  EC~\cite{Dechter:2013:BLV:2540128.2540316}.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

@@

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lllllll}
  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}}& EC&RF/DC & PCFG & Enum

@@

  \\\midrule\multicolumn{7}{c}{\emph{List Processing}}\\\midrule
  \% solved&\textbf{86\%} &84\% &&60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&&1.0s&1.0s&1.1s

  \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
  \% solved&\textbf{75\%} &43\% &&33\%&0\%&4\%\\
  Solve time&  29s&65s&&80s&--&235s

@@

  \\\midrule\multicolumn{7}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  \textbf{24}s& 40s  &28s&31s&55s&29s
  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6

  \\\bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
